---
title: "R Notebook"
output: html_notebook
---


# Text mining homework

```{r}
library(textdata)
library(tidytext)
library(janeaustenr)
library(tidyverse)
```

## Find the most common words in both Pride & Prejudice and Sense & Sensibility.

```{r}
head(prideprejudice, 10)
```
```{r}
head(sensesensibility, 20)
```

```{r}
pride_book <- tibble(
  id = 1:length(prideprejudice),
  text = prideprejudice
) %>% 
  unnest_tokens(word, text)

head(pride_book, 10)
```

```{r}
sense_book <- tibble(
  id = 1:length(sensesensibility),
  text = sensesensibility
) %>% 
  unnest_tokens(word, text)

head(sense_book, 10)
```

```{r}
top_words_pride <- pride_book %>% 
  count(word, sort = TRUE) %>% 
  arrange(desc(n)) %>% 
  head(20) %>% 
  pull(word)

top_words_sense <- sense_book %>% 
  count(word, sort = TRUE) %>% 
  arrange(desc(n)) %>% 
  head(20) %>% 
  pull(word)

pride_sense <- tibble(
  top_words_pride,
  top_words_sense
)

pride_sense
```

## Find the most common words in both Pride & Prejudice and Sense & Sensibility, not including stop words.

```{r}
stop_words_pride <- pride_book %>% 
  anti_join(filter(stop_words, lexicon == "snowball")) %>% 
  count(word, sort = TRUE) %>% 
  arrange(desc(n)) %>% 
  head(20) %>% 
  pull(word)

stop_words_sense <- sense_book %>% 
  anti_join(filter(stop_words, lexicon == "snowball")) %>% 
  count(word, sort = TRUE) %>% 
  arrange(desc(n)) %>% 
  head(20) %>% 
  pull(word)

stop_pride_sense <- tibble(
  stop_words_pride,
  stop_words_sense
)

stop_pride_sense
```


## 3. Find the most common sentiment words in both Pride & Prejudice and Sense & Sensibility.

```{r}
pride_book %>% 
  inner_join(get_sentiments("bing")) %>% 
  group_by(sentiment, word) %>% 
  count(word, sort = TRUE)
```

```{r}
sense_book %>% 
  inner_join(get_sentiments("bing")) %>% 
  group_by(sentiment, word) %>% 
  count(word, sort = TRUE)
```

## Extension: Taking your results above. Can you create a plot which visualises the differences between the books?

```{r}
pride_senti <- pride_book %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(id) %>% 
  summarise(mean_sentence_pride = mean(value))

sense_senti <- sense_book %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(id) %>% 
  summarise(mean_sentence_sense = mean(value))

books_join <- full_join(sense_senti, pride_senti)

```

```{r}
books_join_long <- books_join %>% 
  pivot_longer(cols = !id, names_to = "book", values_to = "mean_sentence") %>% 
  mutate(book = str_remove(book, "mean_sentence_"))
```

```{r}
books_join_long %>% 
  ggplot(aes(x = id, y = mean_sentence))+
  geom_point(size = 1, alpha = 0.5)+
  geom_smooth(aes(colour = book), linewidth = 1)+
  geom_hline(yintercept = 0, colour = "blue", linewidth = 1.3)+
  ylim(-3.5, 4.5)+
  labs(
    x = "Book sentence",
    y = "Mean sentence value",
    title = "Sentiment Analysis of 'Pride & Prejudice' and 'Sense & Sensibility' Books"
  )
```
Not terribly clear, try to calculate the difference between the sentiment values for each book

```{r}
summary(books_join)

books_no_nas <- books_join %>% 
  mutate(mean_sentence_sense = coalesce(mean_sentence_sense, 0),
         mean_sentence_pride = coalesce(mean_sentence_pride, 0))
```

```{r}
books_no_nas %>% 
  mutate(diff = mean_sentence_sense - mean_sentence_pride) %>% 
  select(id, diff) %>% 
  ggplot(aes(x = id, y = diff))+
  geom_point()+
  geom_smooth(linewidth = 3)+
  geom_hline(yintercept = 0, colour = "red", linewidth = 1)+
    labs(
    x = "Book sentence",
    y = "Difference in mean sentence value",
    title = "Difference in Sentiment Analysis of 'Pride & Prejudice' and 'Sense & Sensibility' Books")+
  theme(plot.title = element_text(hjust = 1))
```

