---
title: "Dplyr weekend homework - Week1"
output: html_notebook
---

Load packages and read in data
```{r}
library(tidyverse)
books <- read_csv("books.csv")
```

## Investigate data, are there any missing values?

```{r}
summary(books)
```



```{r}
head(books)
```



```{r}
# Are there any NAs in the data?
books %>% 
  summarise(across(.fns = ~sum(is.na(.x))))
```



# Q1 What languages were books published in and how many for each language? 

Assuming English is the top language, what is the second most popular and what are the top 10 rated books in that language?


```{r}
# what different languages are listed?
books %>% 
  distinct(language_code)
```


```{r}
# group and summarise books by language
books_languages <- books %>% 
  group_by(language_code) %>% 
  summarise(language_count = n()) %>% 
  arrange(desc(language_count))
books_languages
```

```{r}
# what are the top 10 books in Spanish?

books %>%
  filter(language_code == "spa") %>% 
  arrange(desc(average_rating)) %>% 
  head(10)
```

# Q2: When were'Good reads' books published, and how many for each year? Which years had the highest number of published books?

```{r}
# select columns for title, author and year published. Assign to new name
books_year <- books %>%
  select(rowid, bookID, title, authors, average_rating, publication_date, ratings_count)

# change date format from character to date
books_year_new <- books_year %>%  
   mutate(publication_year = as.Date(publication_date, "%m/%d/%Y")) 
```



```{r}
# format publication from YYYY-MM-DD to  show year YYYY
books_year_new <- books_year_new %>% 
  mutate(year = as.numeric(format(publication_year, "%Y")))
```


```{r}
# group the books by publication year, summarise and sort by year with most books published
books_year_new %>% 
  group_by(year) %>% 
  summarise(mean_rating = mean(average_rating),
            count = n()) %>% 
  arrange(desc(count))
```





# Q3: What is the average rating for each publisher

```{r}
books %>% 
  group_by(publisher) %>% 
  summarise(ave_publisher_rating = mean(average_rating),
            books_per_publisher = n()) %>%
  arrange(desc(books_per_publisher))
```

## Q4: What books have the most rating counts?  

```{r}
# combine review counts and text counts into a new column for numbers of individuals responding. Assuming everyone who left a text review also rated. 
books_count_combined <- books %>%
  mutate(ratings_total_individuals = ratings_count - text_reviews_count,
         .after = text_reviews_count) 

# check if any of these are negative numbers, to check assumption is justified. Only 2 cases, so ok to continue
books_count_combined %>% 
  filter(ratings_total_individuals < 0)
```

```{r}
books_count_combined <- books_count_combined %>% 
  arrange(desc(ratings_total_individuals))

head(books_count_combined, 10)
```




# Q5: How many books have an average rating of 5, 4, 3, 2, and 1. 

Group them by the rating and calculate mean and numbers for each. Can we relabel the books with <100 ratings, as these don't have enough ratings

```{r}
# select variables for title, rating and combined_count. 
books_grouped <- books_count_combined %>% 
  select(rowid, bookID, title, average_rating, ratings_total_individuals) 
```



```{r}
# recode books with less than 100 reviews and assign 'new'star' label to group
books_rated_grouped <- books_grouped %>% 
  mutate(
    ratings_grouped = case_when(
      ratings_total_individuals < 50 ~ "Fewer than 50 reviews",
      average_rating > 4.75 ~ "5 star",
      average_rating > 4 ~ "4 star",
      average_rating > 3 ~ "3 star",
      average_rating > 2 ~ "2 star",
      average_rating > 1 ~ "1 star",
      average_rating > 0 ~ "unrated",
      TRUE ~ "other rating")) 
books_rated_grouped %>%      
head(10)
```
Summarise these groups with mean rating and number of books in each group 


```{r}
 # group books by average rating
books_rated_grouped %>% 
  group_by(ratings_grouped) %>% 
  summarise(mean_rating = mean(average_rating),
            book_count = n())
```

# Q6: what were the top 10 rated books published in the 1980s

```{r}
books_1980s <- books_year_new %>% 
  filter(year == c(1980 : 1989)) %>% 
  arrange(desc(average_rating))  

books_1980s %>% 
  head(10)
```


