---
title: "R Notebook"
output: html_notebook
---


# Week 6, day 2 homework - Distributions

```{r}
library(janitor)
library(tidyverse)
```

## Question 1: load data and explore it
```{r}
mobile_data <- read_csv("data/20190928-items.csv") %>% 
  clean_names()
```

```{r}
head(mobile_data)
```

```{r}
glimpse(mobile_data)
```
## Question 2: Find the brand with the highest numbers of phones in the dataset.

```{r}
mobile_data %>% 
  group_by(brand) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
```
## Question 3: For your top brand, plot the distribution of phone ratings as a probability density, overlaying a fitted normal distribution. Do you think the normal distribution provides a good model of these ratings?

Step 1: first calculate the mean and standard deviation for the ratings of Samsung phones

```{r}
mobile_data_stats <- mobile_data %>% 
  filter(brand == "Samsung") %>% 
  summarise(num = n(),
            mean = mean(rating),
            sd = sd(rating))

mobile_data_stats
```

Step 2: plot the distribution of ratings for Samsung phones


```{r}
mobile_dist <- mobile_data %>% 
  filter(brand == "Samsung") %>% 
  ggplot(aes(x = rating))+
  geom_histogram(col = "white", bins = 40)

 mobile_dist
```


Step 2: calculate the probability density and overlay the fitted normal distribution curve

```{r}
mobile_data %>% 
  filter(brand == "Samsung") %>% 
  ggplot(aes(x = rating))+
  geom_histogram(aes(y = after_stat(density)), col = "white", bins = 40)+
  stat_function(
    fun = dnorm,
    args = list(
      mean = mobile_data_stats$mean,
      sd = mobile_data_stats$sd
    ),
    col = "red"
)
```
The distribution curve fits the data reasonably well, the mean and sd correspond to the calculated values earlier. It looks slightly left skewed, due to some very low ratings. 


Question 4: We can be more quantitative in comparing the distribution of top brand ratings with a normal distribution. You calculated the mean() and sd() of your distribution in an earlier question. Use these values to calculate the proportions of ratings within one-, two- and three standard deviations of the mean. Compare these proportions with those you expect for a normal distribution.


step 1: 

Calculate z-scores for the salary
```{r}
samsung_data <- samsung_data %>% 
  mutate(z_score = scale(rating))
```


Filter the ratings with a z score of +/- 1SD (0.6847859)

```{r}
samsung_data %>% 
  filter(z_score >= -0.685,
        z_score <= 0.685) %>% 
  summarise(rating_total = n(), 
            proportion_total = (rating_total/720) * 100)
```

```{r}
samsung_data %>% 
  nrow()
```



Step 2: 
Calculate the proportion of rating within 1SD
```{r}
samsung_data %>% 
  filter(z_score >= -0.685,
        z_score <= 0.685) %>% 
  summarise(rating_total = n(), 
            proportion_total = rating_total/792 * 100)
```
Calculate the proportion of rating within 2SD
```{r}
samsung_data %>% 
  filter(z_score >= (-0.685 * 2),
        z_score <= (0.685 * 2)) %>% 
  summarise(rating_total = n(), 
            proportion_total = rating_total/792 * 100)
```
Calculate the proportion of rating within 3SD

```{r}
samsung_data %>% 
  filter(z_score >= (-0.685 * 3),
        z_score <= (0.685 * 3)) %>% 
  summarise(rating_total = n(), 
            proportion_total = rating_total/792 * 100)
```























