---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(broom)
library(factoextra)
```
# Task 1: Do some exploratory analysis on the data. Comment on findings.
```{r}
mall_customers <- read_csv("data/mall_customers.csv") %>% 
  clean_names()
```


```{r}
skimr::skim(mall_customers)
```

```{r}
summary(mall_customers)
```

```{r}
mall_customers %>% 
  ggplot(aes(x = annual_income_k, y = spending_score_1_100, colour = gender))+
  geom_point()
```




```{r}
mall_customers %>% 
  ggplot(aes(x = age, y = spending_score_1_100, colour = gender))+
  geom_point()
```
```{r}
mall_customers %>% 
  ggplot(aes(x = age, y = annual_income_k, colour = gender))+
  geom_point()
```

From the graph correlating spending score with annual income, it looks initially like there are 5 groups, 4 arranged around a central group. 

Looks like there are group of incomes arranged in a line at £54k and £78k, could be a default income? 

Age and gender seem to be fairly evenly distributed, no discernable patter to these variables. 

The age vs spending_score is interesting as the lower ages are spending the highest amounts, whereas older age groups have lower spending scores. 


# Task 2: We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Select an appropriate value for k and perform k-means clustering to find if there are meaningful clusters in the data to target the customer



```{r}
# scale the data so that the higher numbers don't disproportionally influence the mean
mall_cust_scaled <- mall_customers %>% 
  select(annual_income_k, spending_score_1_100) %>% 
  mutate(across(everything(), 
         .fns = scale))
```



```{r}
# create data for around a range of clusters amounts that we want to check
min_k <- 2
max_k <- 10

k_clusters <- tibble(k = min_k:max_k) %>% 
  mutate(kclust = map(k, ~kmeans(mall_cust_scaled, centers = .x, nstart = 25)),
         tidied = map(kclust, tidy),
         glance = map(kclust, glance),
         augmented = map(kclust, augment, mall_customers))

k_clusters
```


```{r}
# Unnest glance to choose the number of k
clusterings <- k_clusters %>% 
  unnest(glance) 

clusterings
```
To choose the most appropriate value for k we will first plot the sum of squares (withinss), which is a single measure of how well compacted the k clusters are. This is called the elbow method and we are looking for the value were the angle is most strongly influenced by the next point

```{r}
# elbow method
clusterings %>% 
  ggplot(aes(x = k, 
             y = tot.withinss))+
  geom_point() +
  geom_line()+
  scale_x_continuous(breaks = seq(1, 20, 1))
```


Second plot will be the "silhouette" method, which is similar to a 'goodness of fit'. The higher the better, and greater than 0.5 is good
```{r}
fviz_nbclust(mall_cust_scaled, 
             kmeans,
             method = "silhouette",
             nstart = 25)
```
Last method is called gap_stat
```{r}
fviz_nbclust(mall_cust_scaled, 
             kmeans,
             method = "gap_stat",
             nstart = 25)
```
Elbow method suggests k = 5
Silhouette method suggests k = 5
Gap statistic method suggests k = 1, however the highest value is seen at 5 clusters. 

Two of the three methods suggest that 5 is the optimal number of clusters, so will proceed with 5.

# Task 3: Visualise the clustering for your chosen value of k.


```{r}
clusterings %>% 
  unnest(augmented) %>% 
  filter(k == 5) %>% 
  ggplot(aes(y = spending_score_1_100, x = annual_income_k, 
             colour = .cluster))+
  geom_point()
```
# Task 4: Do you think the clustering seems a good fit for this data?

Yes.Data exploration originally identified 5 groups, which have been assigned in this figure. 


# Task 5: Comment on the attributes of one or two of the clusters (maybe even give them an appropriate label if you like - similar to section 4.1 of the ‘Segmentation & clustering intro’ lesson where we saw labels such as “Super Fans” and “Tag-Alongs”).


```{r}
clusterings %>% 
  unnest(augmented) %>% 
  filter(k == 5) %>% 
  group_by(.cluster) %>% 
  summarise(ave_income = mean(annual_income_k),
            ave_age = mean(age),
            ave_spending_score = mean(spending_score_1_100))
```
Could assign group labels dependent on age and earnings

Group 1: >40s / middle earners / middle spenders
Group 2: >40s / high earners/ low spenders
Group 3: <40s / low earners / high spenders
Group 4: >40s / low earners / low spenders
Group 5: <40s / high earners / high spenders

Might be worth targetting customers who fit group2 demographics as they have disposable income and don't spend my in this shop. 
