---
title: "Untitled"
output:
  html_document: default
  pdf_document: default
date: "2023-03-28"
---

```{r}
library(tidyverse)
tweets <- read_csv("code_clan_tweets.csv")
dim(tweets) # 234 rows,27 columns
```


```{r}
names(tweets)
```
Question 2: Find the total number of favourites (stored in favorite_count) that CodeClan tweets have got. Don’t add a tweet’s favorite_count to the total when the tweet was quoted (see the is_quote variable).

```{r}
favourites_total <- tweets %>% 
  filter(is_quote == FALSE) 
```


```{r}
favourites_total %>% 
  summarise(favorite_total = sum(favorite_count))
```



Question 3: Summarise the mean number of retweets (retweet_count), split by the type of platform that was used to tweet (source). Again, omit quoted tweets from the mean.

```{r}
tweets %>% 
  distinct(source)
```



```{r}
retweet_count_mean <- tweets %>% 
  filter(is_quote == FALSE) %>% 
  group_by(source) %>% 
  summarise(mean_retweets = mean(retweet_count))
retweet_count_mean
```

Question 4.
Count the total number of likes (i.e. total of favorite_count), split by media type, and arrange them from most likes to least. Assume that any tweets without a listed media type are of type “text”.

step 1: convert NAs in Media to "Text", then group and summarise
```{r}
total_likes <- tweets %>%
  mutate(media_type = coalesce(media_type, "text")) %>%
  group_by(media_type) %>% 
  summarise(total_likes = sum(favorite_count)) %>% 
  arrange(desc(total_likes))
  
total_likes
```

Question 5.
Find the mean number of characters that a CodeClan tweet contains. You can either find a way to count the text in the text variable, or use the variable display_text_width which contains a count. Checking both would be a good idea if possible to see if they match (but it’s not essential).


```{r}
# count the length of string using 'str_length' then check if this is equal to display text width
tweet_length <- tweets %>% 
  mutate(tweet_length = str_length(text), .after = text,
         length_check = (tweet_length == display_text_width))
```



Question 6.
The code_clan_info.csv data file contains status_url along with other info. Load this in, and join it to the code_clan_tweets tibble, so that you have a status_url for each tweet. Decide which variable to join the tibbles on.

```{r}
info <- read_csv("code_clan_info.csv")
```

```{r}
names(info)
```
```{r}
tweet_join_info <- 
  full_join(tweets, info, by = "tweet_id")

names(tweet_join_info)
```
```{r}
dim(tweet_join_info)
```

Question 7.
From your new joined data, create a new tibble codeclan_hashtags containing only tweet_id and hashtags, then convert hashtags into lowercase for analysis. Keep only those tweets with hashtags.


```{r}
codeclan_hashtags <- tweet_join_info %>% 
  select(tweet_id, hashtags) %>% 
  mutate(hashtags = str_to_lower(hashtags)) %>% 
  filter(!is.na(hashtags))
head(codeclan_hashtags)
```

Extension: Question 8
Some tweets have more than one hashtag, denoted by the c( symbols at the start of the string. Use the str_sub() and str_detect() functions and your codeclan_hashtags tibble from above to find all the cases in which the hashtag string begins with charactersc(.
























