---
title: "Titanic Homework - week6 day 1"
output: html_notebook
---
load libraries
```{r}
library(tidyverse)
library(gtools)
library(janitor)
source("prob.R")
```


read in data
```{r}
titanic <- read_csv("data/titanic.csv")
```

Calculate the conditional probabilities prob(survived∣crew)
 and prob(survived∣passenger)
 to settle this dispute.
 
 
 Step 1. create a new column 'crew' with TRUE/FALSE 

```{r}
titanic %>% 
  distinct(class)
```

```{r}
titanic %>% 
  group_by(class) %>% 
  count()
```


 
```{r}
crew <- titanic %>% 
  mutate(crew = ifelse(class %in% c("engineering crew", "victualling crew", "restaurant staff", "deck crew"), "TRUE", FALSE))
```


Step 2: calculate the proportion of crew that survived
```{r}
crew %>% 
  filter(crew == TRUE) %>% 
  summarise(prob = sum(survived == "yes")/ n())
```

 
Calculate proportion of passengers to survive

```{r}
crew %>% 
  filter(crew == FALSE) %>% 
  summarise(prob = sum(survived == "yes")/ n())
```
THerefore a higher proportion of passengers survived the Titanic sinking. 



## Question 1: If I roll a six-sided dice once, what is the probability of getting a number less than 5?

Prob(event) = number_outcomes / number_all_outcomes 

Prob(rolling_4) = 1, 2, 3, 4, / 1, 2, 3, 4, 5, 6)
Prob(rolling_4) = 4 / 6 = 0.667


## Question 2: How many outcomes are there in the sample space of an experiment in which we toss three dice?

```{r}
# sample space for roll of 3 dice
s_three_dice <- rolldie(3)
s_three_dice %>% 
  nrow()
```
Question 3: In an experiment where we roll four dice:

- how many outcomes are in the event where the total of the four dice equals 5, 6 or 7)?
n(total_5_6or7) = 34

```{r}
s_four_dice <- rolldie(4)

s_four_dice %>% 
  mutate(total = X1 + X2 + X3 + X4) %>% 
  filter(total == 5 |
         total == 6 |
         total == 7)
```


- what is the probability of this event?
```{r}
s_four_dice %>% 
  nrow()
```
Probability of rolling 4 dice with total 5, 6, or 7 =
Prob(rolling_5_6or7) = n(total_5_6or7) / n(all_outcomes) = 34 / 1296 = 0.0262




- what assumptions do we apply in answering part ii?
That all numbers are weighted equally 
 
## Question 4: Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?


```{r}
heights <- c("Al", "Bobby", "Charlotte", "David", "Emily")

length(heights)
# n = length(candidates)
# r = 5
# v = candidates
heights_order <- as_tibble(
  permutations(n = length(heights),
               r = 5,
               v = heights,
               repeats.allowed = FALSE))

nrow(heights_order)

```
There are 120 possible combinations, with only one being the correct order. Therefore
prob = 1 / 120
= 0.0083


# Question 5: I roll six dice:

- what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right?

```{r}
dice <- c(1, 2, 3, 4, 5, 6)

dice_order <- as_tibble(
  permutations(n = length(dice),
               r = 6,
               v = dice,
               repeats.allowed = TRUE)
) 
nrow(dice_order)

```


```{r}
dice_order %>% 
  filter(V1 == 6,
         V2 == 5, 
         V3 == 4,
         V4 == 3,
         V5 == 2,
         V6 == 1)
```
Prob(6_1ordered) = nrow(6,5,4,3,2,1) / nrow(all)
                    1 / 46656
                    0.0000214


- what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order?


```{r}
dice_comb <- as_tibble(
  combinations(n = length(dice),
               r = 6,
               v = dice,
               repeats.allowed = TRUE)
) 
nrow(dice_comb)
```
Prob(6:1) = nrow(6-1) / nrow(all)
                    1 / 462
                    0.00216
                    
                    
                    
# Question 6: If I draw a single card from a 52-card deck, and it is black, what is the probability that its suit is spades?

conditional probability where:
event A: draw spades
event B: draw black card



Step 1: calculate prob(A n B)
prob(A n B) = num(A n B) / num(all)
            = 13        / 52
            0.25


Step 2: Calculate probability of spades|black card

p(spades|black_card) = prob(A n B) / prob(black_card)
prob(spades|black) = (0.25) / (0.5)
                  = 0.5
                  

Therefore, probability of blackcard and spades is 0.5
